### 2021-11-19 article.tex
rearrange figures and write more on double slit experiment

\documentclass{article}
\pdfsuppresswarningpagegroup=1

\author{Chinmaya Nagpal}
\title{May I Ask Where You Have Been, Photon?}
\date{18 November 2021}

\usepackage{svg,subfig,framed}
\usepackage[a4paper, total={6in, 8in}]{geometry}

\newcommand{\Q}{\bfseries Q: }
\newcommand{\A}{\par\textbf{A:} \normalfont}

\begin{document}

\maketitle

\begin{figure}[ht]
	\subfloat[Standard 
	setup]{\fbox{\includesvg[width=0.5\linewidth]{figures/double-slit.svg}}\label{fig:uncollimated}}
	\hspace*{\fill}
	\subfloat["Which-way" experiment
	]{\fbox{\includesvg[width=0.5\linewidth]{figures/collimated-double-slit.svg}}\label{fig:collimated}}
	\caption{Young's double slit interference experiment}
	\label{fig:double-slit}
\end{figure}

Quantum physics describes physical interactions that happen at the atomic and
sub-atomic levels, and underpins the foundation for most of modern technology.
However, physicists are still unsure how to use standard quantum theory to 
answer the following question: what path did a photon take through an 
interferometer before quantum measurement was performed?  A recent experiment 
conducted by A. Danan and colleagues at Tel Aviv University in Israel sheds 
light on how difficult it is to answer this question in any but the simplest of 
situations.

In classical physics, a particle travelling between two points has a single
path, which can be calculated by specifying its state at either endpoint. On 
the other hand, in quantum physics, the trajectory a particle takes between two 
measurements is a weighted sum, or superposition, of every path it could have 
taken between the two measurements. This gives rise to the phenomenon of a 
particle interfering with itself before being measured, as demonstrated by 
experiments like the single-particle version of Young's double slit experiment.

In the classical double slit experiment, a beam of light is passed through two 
closely spaced slits and projected onto a screen. Because of the wave nature of 
light, different distances travelled by different wavefronts give rise to phase 
differences at the screen, causing an interference pattern to appear, as 
illustrated in figure~\ref{fig:uncollimated}. This is similar to interference 
between ripples in a pond, creating zones of zero displacement where crests and 
troughs overlap.

In the modern version of this experiment, the continuous light source is 
replaced by a light source emitting individual particles, and a photosensitive 
plate is used as the screen. Over time, the same interference pattern appears 
on the photosensitive plate as in the classical experiment. This implies that 
individual particles undergo wave-like interference \textit{with themselves}, 
begging the question:

\begin{framed}
\Q Where have you been, photon?

\A Because the particle underwent wave-like interference with itself, it can be 
	said that it took \textit{all possible paths} between the light source 
	and the screen.  However, because it was not observed at any point 
	between the source and the screen, a common position is to claim that
	\textit{one cannot say} where the photon has been.
\end{framed}

One way to try to ascertain where each photon is coming from (in a so-called 
"which-way" experiment) is to insert a lens after each slit, as shown in 
figure~\ref{fig:collimated}.  Each lens only collimates the photons that arrive 
through the corresponding slit, causing the interference pattern to be replaced 
by two distinct spots. It would then seem safe to assume that a photon detected 
in the a spot on the screen must have travelled through the corresponding slit.  
It does not seem possible that a photon detected in, say, the upper spot would 
have gone through the lower lens. It is this seemingly safe assertion that is 
called into question by the Tel-Aviv experiment.

In their experiment, Danan and collagues use the Mach-Zehnder interferometer, 
used by Danan and colleagues in their experiment.  The Mach Zehnder 
interferometer consists of a light source and a detector along with two beam 
splitters (basically half-silvered mirrors) and two mirrors, arranged as shown 
in figure~\ref{fig:interferometer1}. In the classical version of the 
experiment, the light source emits a continuous beam of light, which partly 
passes through the the first beam splitter, and is then reflected by the two 
mirrors to meet at the second beam splitter. Reflection at an interface between 
vacuum and the silvered side of a mirror causes a phase change through a purely 
classical phenomenon, which means that the beams coming out of the second beam 
splitter combine totally constructively on one side, and totally destructively 
on the other, in analogy to the bright and dark bands in the classical double 
slit experiment. Extending the analogy to the double-slit experiment, if the 
light source in the Mach-Zehnder interferometer is tuned to a low-enough 
intensity that single photons emitted can be resolved by the detector, it is 
found that the photon arrives at the detector 100\% of the time, demonstrating 
the phenomenon of a particle interfering with itself. Of course, if we remove 
the second beam splitter, there is a 50\% chance that the photon will show up 
in our detector, and a 50\% chance that it will come out of the other branch of 
the interferometer.

\begin{figure}
	\includesvg[width=\linewidth]{figures/interferometer1.svg}
	\caption{Simple Mach Zehnder interferometer}
	\label{fig:interferometer1}
\end{figure}

Again, we can answer titular question using "common sense" approach amateur 
science enthusiasts and undergraduate physics students are familiar with:

\medskip\Q May I ask where you have been, photon?

\A The photon's trajectory in the interferometer is a superposition of the 
classical trajectories through the upper and lower arms of the 
interferometer.\medskip

The Tel-Aviv experiment starts off with a modified version of the Mach-Zehnder 
interferometer, in which each mirror oscillates vertically with a unique 
frequency.  Additionally, instead of a simple photodetector, a stacked array of 
detectors is used, making it possible to monitor the frequency spectrum of the 
detected light wave. As expected, the power spectrum shows equal contributions 
at frequencies $f_A$ and $f_B$ associated with mirrors $A$ and $B$.

Whilst this does not directly answer our question, since the photon source 
emits numerous photons, which means that different ones could have caused the 
peaks at $f_A$ and $f_B$, it at least confirms that some of the photons have 
been at each mirror.

In the second variation of the experiment, the same steps are repeated, but 
with the second beam splitter removed. This time, only the peak at $f_B$ 
remains, again confirming our interpretation of the Mach-Zehnder interferometer 
experiment.

Next, a larger interferometer is built, as illustrated in figure 
\ref{fig:danan-2a}. There is a nested Mach-Zehnder interferometer in the upper 
arm, and the beam intensity is now split 1:2 between the upper and the lower 
arm, so the ratio of intensities at the vibrating mirrors $A$, $B$, and $C$ is 
1:1:1.  Again, the power spectrum shows what we would expect: peaks at all 
frequencies, with larger peaks at $f_E$ and $f_F$.

So far, the results of this experiment have confirmed our intuitive, "common 
sense" approach to accounting for a quantum particle's trajectory as a sum of 
all possible paths. The surprising result from this experiment is obtained when 
the following changes are made, as illustrated in figure \ref{fig:danan-2b}:
either the second beam splitter of the mini-Mach Zehnder interferometer is 
flipped, or mirror $B$ is moved to slightly adjust the path length, causing 
destructive interference of the light propagating towards $F$.

\begin{figure}
	\subfloat[]{\includesvg[width=0.5\linewidth]{figures/danan-2a.svg}\label{fig:danan-2a}}
	\subfloat[]{\includesvg[width=0.5\linewidth]{figures/danan-2b.svg}\label{fig:danan-2b}}
	\newline
	\hspace*{\fill}
	\subfloat[]{\includesvg[width=0.5\linewidth]{figures/danan-2c.svg}\label{fig:danan-2c}}
	\hspace*{\fill}
	\caption{Nested Mach-Zehnder interferometer}
	\label{fig:danan}
\end{figure}

\end{document}